---
title: "Calculating food web and community metrics"
author: "Laurene Pecuchet"
date: '2022-06-08'
output:
  word_document: 
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(NetIndices)
library(reshape2)
library(igraph)
library(vegan)
library(tidyverse)
library(dggridR)
library(gridExtra)
library(Hmisc)


#Construct a global grid with cells approximately 1000 miles across
dggs          <- dgconstruct(res=7, metric=FALSE, resround='down')


```

## Demersal community and Food web (data preparation)

### Assembling community per grid cell

For each grid cell we will calculate community and food web metrics
based on the fish and megabenthos trophospecies presence and biomass in
the bottom-trawl survey (communities). The first step is to do sampling
effort standardization by assembling communities per grid cell from a
random draw of 10 sampled hauls made in the period 2009-2017, iterated
100 times (i.e., spatial snapshot without temporal consideration). 
Raw data is not available for this calculation due to data restrictions.

```{r Metacommunities, eval=F, include=F}
com_fw<-communities 
head(communities)#"communities" is the data.frame of randomly assembled communities
com_fw<-merge(com_fw, name_survey[,c("ABBREV_SIMPL","ScientificName")], by.x="variable", by.y="ScientificName")
com_fw<-com_fw %>% group_by(cell, ABBREV_SIMPL, iter) %>% dplyr::summarise(biomass=sum(bio, na.rm=T))
head(com_fw) #"com_fw" is the summed biomass per taxon per iteration and cell, with a standardized name matching between the surveys and the food web. This is the "community composition" dataframe that is used to calculate spatial community and food web metrics.

save(com_fw, file="Results/metacommunities1000_fw.Rdata")
```

## Calculate metrics: Communities and Food web

### Traits: size, generality, trophic level, omnivory

```{r, Traits spatially fixed}
TS_traits<-read.table("Data/traits_prop.csv", header=T, sep=";")
TS_traits[!TS_traits$Group=="Fish",]$Group<-"Benthic.invertebrates"
traits_m<-TS_traits %>% group_by(ABBREV_SIMPL) %>% dplyr::summarise(Size=mean(continuous_cm, na.rm=T),
                                                                 Group=unique(Group))
traits_feed<-data.frame(ABBREV_SIMPL=rownames(TrophInd(fw_new)), TrophInd(fw_new), Gen=colSums(fw_new))

traits_m<-merge(traits_feed, traits_m, by="ABBREV_SIMPL", all.x=T)
traits_m$Size.log<-log(traits_m$Size+1)
hist(traits_m$Size, main="Size (cm)")
hist(traits_m$Size.log, main="Size log-transformed log(cm)")

species_list<-merge(trophind_qual, traits_m, by.x="species", by.y="ABBREV_SIMPL")
#write.table(species_list, "Specieslist_traits.csv", sep=";")

```

How well are the bentho-demersal communities represented by the food web? We calculate and map per grid cell the % of the individuals sampled the were included in the food web. (Raw data is not available for this calculation due to data restrictions)

```{r, eval=F}
load("Results/com_data1000.Rdata")
head(communities)

communities<-merge(communities, name_survey[,c("ABBREV_SIMPL","ScientificName")], by.x="variable", by.y="ScientificName", all.x=T)

com.NA<-communities %>% filter(is.na(ABBREV_SIMPL)) %>% group_by(cell, ABBREV_SIMPL, with.dml, iter) %>% dplyr::summarise(bio.NA=sum(bio, na.rm=T))

com.Info<-communities %>% filter(!is.na(ABBREV_SIMPL)) %>% group_by(cell, with.dml, iter) %>% dplyr::summarise(bio.Info=sum(bio, na.rm=T))

com.NA<-left_join(com.Info,com.NA)
com.NA$percInfo<-(1-com.NA$bio.NA/(com.NA$bio.Info+com.NA$bio.NA))*100

com.NA<-com.NA%>% group_by(cell, with.dml) %>%  dplyr::summarise(percInfo5=ci(percInfo)$CI_low,percInfo95=ci(percInfo)$CI_high,
  percInfo=median(percInfo, na.rm=T))

com.NA[is.na(com.NA$percInfo),]$percInfo<-100
## If NA = 100% sampled individuals included in the food web

com.NA.F<-com.NA %>% filter(with.dml==1)
summary(com.NA.F)

#Get the grid cell boundaries
grid0          <- dgcellstogrid(dggs,com.NA.F$cell,frame=TRUE,wrapcells=TRUE)
grid          <- merge(grid0,com.NA.F,by.x="cell",by.y="cell")

p1<-plot.fw(grid$percInfo, "")+ ggtitle("a) Fish")

com.NA.B<-com.NA %>% filter(with.dml==0)
summary(com.NA.B)

#Get the grid cell boundaries
grid0          <- dgcellstogrid(dggs,com.NA.B$cell,frame=TRUE,wrapcells=TRUE)
grid1          <- merge(grid0,com.NA.B,by.x="cell",by.y="cell")

p2<-plot.fw(grid1$percInfo, "")+ ggtitle("b) Benthos")

tiff("Results/PercIndiv.tiff", width=20, height=10, units="cm", res=400)
grid.arrange(p1, p2, ncol=2)
dev.off()

```

Biomass is log transformed before calculating CWM feeding and size
traits Biomass is not log-transformed for calculating the Biomass
evenness and the total biomass The trait "Size" is also log-transformed
due to right-skewness

We calculate multiple community and food web metrics for each assembled
communities per grid cell and iteration (=100 assembled communities \*
60 grid cell). At the community level, we calculate biomass-weighted
indices of: total biomass, biomass eveness, biomass ratio
fish/invertebrates and community mean and variance of body size, trophic
level, omnivory and generality. At the food web level, we calculate
unweighted (i.e., presence-based) food web properties of number of taxa,
number of links, link density, connectance, modularity, mean chain
length, nestedness

Species' communities data is not available for this calculation due to data restrictions.

```{r, eval=FALSE}
is.nan.data.frame <- function(x) do.call(cbind, lapply(x, is.nan))

load("Results/metacommunities1000_fw.Rdata")

com_fw$bio.log<-log(com_fw$biomass+1)
com_fw<-com_fw %>% dplyr::filter(!is.na(ABBREV_SIMPL))

#fw.sp<-data.frame()

for(s in unique(com_fw$cell)){ 
  print(s) 
    for (i in 1:100){ # here stop at 100 iterations only to speed the calculations, can change to 1000
      print(i)
      
com<-subset(com_fw, cell==s&iter==i)

### Community-level indices

## Biomass metrics, relevant for ecosystem functions
# Biomass evenness 
bioE=vegan::diversity(com$biomass, index="simpson") # Biomass evenness using the simpson index
# Total biomass
totbio<-sum(com$biomass, na.rm=T)
# Proportion fish to benthos
com_tr<-left_join(com,traits_m)
BF<-com_tr %>% group_by(Group) %>% dplyr::summarise(bio=sum(biomass, na.rm = T)) %>% pivot_wider(names_from = Group, values_from = bio)
FB<-BF$Fish/BF$Benthic.invertebrates

## Trait metrics (Community weighted mean and variance of traits) -> log transformed, relevant for community composition and diversity
# Trophic level
TL.var<-sqrt(wtd.var(as.vector(com_tr$TL), weights= com_tr$bio.log/sum(com_tr$bio.log),method="ML", na.rm=T, normwt=T))
TL<- wtd.mean(as.vector(com_tr$TL), weights= com_tr$bio.log/sum(com_tr$bio.log))
# Omnivory
OI.var<-sqrt(wtd.var(as.vector(com_tr$OI), weights= com_tr$bio.log/sum(com_tr$bio.log),method="ML", na.rm=T, normwt=T))
OI<- wtd.mean(as.vector(com_tr$OI), weights= com_tr$bio.log/sum(com_tr$bio.log))
# Generality
Gen.var<-sqrt(wtd.var(as.vector(com_tr$Gen), weights= com_tr$bio.log/sum(com_tr$bio.log),method="ML", na.rm=T, normwt=T))
Gen<- wtd.mean(as.vector(com_tr$Gen), weights= com_tr$bio.log/sum(com_tr$bio.log))
# Body size
Sizelog.var<-sqrt(wtd.var(as.vector(com_tr$Size.log), weights= com_tr$bio.log/sum(com_tr$bio.log),method="ML", na.rm=T, normwt=T))
Sizelog<- wtd.mean(as.vector(com_tr$Size.log), weights= com_tr$bio.log/sum(com_tr$bio.log))

### Food web properties
      if(dim(com)[1]>0)
      {
       ## Obtain the local food web from the metaweb
       fw<-fw_new[rownames(fw_new)%in%c(as.character(com$ABBREV_SIMPL),"Zooplankton","Phytoplankton","DET_IND"),]
       fw<-fw[,colnames(fw)%in%c(as.character(com$ABBREV_SIMPL),"Zooplankton","Phytoplankton","DET_IND")]

      ### Get rid of unconnected species (species with neither preys or predators)
      el<-intersect(rownames(fw[rowSums(fw)==0,]),colnames(fw[,colSums(fw)==0]))
      fw<-fw[!rownames(fw)%in%el,]
      fw<-fw[,!colnames(fw)%in%el]

pfw<-fw
pfw[pfw>0]<-1 # Presence based indices, all positive biomass>0 become "present" = 1
# Number of species
Nspp<-dim(pfw)[1]
# Number of links
Nlinks<-sum(pfw)
# Link density
LD<-Nlinks/Nspp

# Connectance : nb links/(nb nodes^2)
Conn<-sum(pfw)/dim(pfw)[1]^2

# Nestedness
nested<-nestednodf(pfw)$statistic[3]

# Modularity
fwig<-graph.adjacency(as.matrix(pfw))       # igraph formatted foodwebs
spingc <- tryCatch(spinglass.community(fwig), error=function(err) NA) # modularity foodweb
modul  <- tryCatch(spingc$modularity, error=function(err) NA) # modularity as defined by Newman and Girvan

# Mean shortest path
meanChain<- mean_distance(fwig)

fw.sp<-rbind(fw.sp,data.frame(cell=s, iter=i, nodes=Nspp, links=Nlinks, Connect=Conn, nested=nested,  modul=modul,  meanChain=meanChain,  Sizelog.sd=Sizelog.var, Sizelog=Sizelog,TL.sd=TL.var, TL=TL, OI.sd=OI.var, OI=OI,LD=LD, Gen.sd=Gen.var, Gen=Gen, bioE=bioE, totbio=totbio, BF, FB=FB))

}
  }
} 

save(fw.sp, file="Results/ComFW.spatial.Rdata")

```
